---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions,echo=TRUE}
library(knitr)
opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data
```{r LoadingProcessingData,cache=TRUE}
unzip("activity.zip")
activity <- read.csv("activity.csv",na.strings="NA",stringsAsFactors = FALSE)
str(activity)
```


## What is mean total number of steps taken per day?
```{r MeanTotalStepsPerDay,cache=TRUE}
steps_per_day <- aggregate(steps~date,data=activity,sum)

require(ggplot2)
library(ggplot2)
ggplot(steps_per_day,aes(steps,fill=date)) +
geom_histogram(binwidth = 53)

na_mean_steps_per_day <- as.integer(mean(steps_per_day$steps))
na_median_steps_per_day <- median(steps_per_day$steps)
```
Mean of the total number of steps taken per day: `r na_mean_steps_per_day`  
Median of the total number of steps taken per day: `r na_median_steps_per_day`  


## What is the average daily activity pattern?
```{r AverageDailyActivityPattern,cache=TRUE}
avg_steps_per_day <- aggregate(steps~interval,data=activity,mean)

ggplot(avg_steps_per_day, aes(interval, steps)) + 
        geom_line()

interval_max_steps <- avg_steps_per_day[avg_steps_per_day$steps==max(avg_steps_per_day$steps),]
```

Interval: `r interval_max_steps$interval`, on average across all the days in the dataset, contains the maximum number of steps: `r interval_max_steps$steps`


## Imputing missing values
```{r ImputMissingValues,cache=TRUE}
na_rows <- nrow(activity[is.na(activity),])

# Used avg_steps_per_day for estimated values of na steps
activity_tidy <- activity
activity_tidy[is.na(activity_tidy)==TRUE] <- as.integer(avg_steps_per_day$steps)

steps_per_day <- aggregate(steps~date,data=activity_tidy,sum)

ggplot(steps_per_day,aes(steps,fill=date)) +
geom_histogram(binwidth = 53)

mean_steps_per_day <- as.integer(mean(steps_per_day$steps))
median_steps_per_day <- median(steps_per_day$steps)
```


Total number of missing values in the dataset (i.e. the total number of rows with NAs): `r na_rows`

Mean of the total number of steps taken per day: `r mean_steps_per_day`  
Median of the total number of steps taken per day: `r median_steps_per_day`  

These values differ from the estimates from the first part of the assignment, an overall increase of max count with a left shift of its step indicator

Imputing missing data on the estimates of the total daily number of steps impacts:

Means went from `r na_mean_steps_per_day` to `r mean_steps_per_day`

Medians went from `r na_median_steps_per_day` to `r median_steps_per_day`

Histograms went from a max count of 3 at appoximately 15000 steps to a max count of 8 at slightly above 10000 steps


## Are there differences in activity patterns between weekdays and weekends?
```{r DifferenceActivityPatternsWeekdaysWeekends,cache=TRUE}
activity_tidy <- transform(activity_tidy,day=factor(ifelse(weekdays(as.Date(activity_tidy$date), abbreviate = FALSE) %in% c("Saturday","Sunday"),"weekend","weekday")))

avg_steps_per_day <- aggregate(interval~steps +day,data=activity_tidy,mean)

ggplot(avg_steps_per_day, aes(interval, steps)) + 
        geom_line() +
        facet_wrap(~day,ncol = 1)
```
